---
title: "2x2x2 Rubik's cube"
author: "Manu"
date: "16 February 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 2x2x2 Rubik's cube
This Rubik's cube has 8 small cubes. We give them a number like this:


```{r cube-def}
cat('Front face \r\n')
matrix(1:4, nrow = 2, byrow = T)
cat('Back face \r\n')
matrix(5:8, nrow = 2, byrow = T)

```

## Moves

Solving a Rubik's cube consist of turning the Rubik's cube in different ways.
Each way to transform the Rubik's cube is called a move. We can see that there
are 6 basic moves: spin top, spin bottom, topple left, topple right, wash front and wash back.

```{r basic-moves}
bg <- cyc(c(1, 5, 7, 3)) %>% as.cycle
bd <- cyc(c(2, 6, 8, 4)) %>% as.cycle
th <- cyc(c(1, 2, 6, 5)) %>% as.cycle
tb <- cyc(c(3, 4, 8, 7)) %>% as.cycle
wf <- cyc(c(1, 2, 4, 3)) %>% as.cycle
wb <- cyc(c(5, 6, 8, 7)) %>% as.cycle
cat(' Toggle left', as.character(bg), '\r\n',
    'Toggle right', as.character(bd), '\r\n',
    'Spin top', as.character(th), '\r\n',
    'Spin bottom', as.character(tb), '\r\n',
    'Wash front', as.character(wf), '\r\n',
    'Wash back', as.character(wb))

```

## Invariants

Rotating the whole cube left, right, up or down intuitively doesn't change
the problem. H

Let r be an element of R that maps to the identity by the group action. 
Conjugation of a basic move by a global rotation yields a basic move, or the inverse of a basic move.

```{r commutativity}
gen <- c(bg, bd, th, tb, wf, wb)
rot_up * gen / rot_up == c(bg, bd, wf^-1, wb^-1, tb, th)
rot_right * gen / rot_right == c(wb, wf, th, tb, bg^-1, bd^-1)
rot_cck * gen / rot_cck == c(tb^-1, th^-1, bg, bd, wf, wb)



```

