---
title: "C(2)xC(2)xC(2)"
author: "Manu"
date: "January 16, 2018"
output: html_document
---

```{r setup, include=FALSE}
library(permutations)
source('utils.R')
knitr::opts_chunk$set(echo = TRUE)
```

# C(2)xC(2)xC(2)

THis group has 8 elements.
It's normal decomposition is $C(1) \lhd C(2) \lhd C(2)^2 \lhd C(2)^3$

The composition factors are 3 $C(2)$

# Beat Action

|Beat       | 1 | 2 | 3 | 4 | 5 | 6 | 7 | 8 |
|-----------|---|---|---|---|---|---|---|---|
|e          | 1 | 2 | 3 | 4 | 5 | 6 | 7 | 8 |
|(1, 0, 0)  | 5 | 6 | 7 | 8 | 1 | 2 | 3 | 4 |
|(0, 1, 0)  | 3 | 4 | 1 | 2 | 7 | 8 | 5 | 6 |
|(0, 0, 1)  | 2 | 1 | 4 | 3 | 6 | 5 | 8 | 7 |
|(1, 1, 0)  | 7 | 8 | 5 | 6 | 3 | 4 | 1 | 2 |
|(1, 0, 1)  | 6 | 5 | 8 | 7 | 2 | 1 | 4 | 3 |
|(0, 1, 1)  | 4 | 3 | 2 | 1 | 8 | 7 | 6 | 5 |
|(1, 1, 1)  | 8 | 7 | 6 | 5 | 4 | 3 | 2 | 1 |

We check that it is indeed an action
```{r}
g100 <- as.word(c(5:8, 1:4))
g010 <- as.word(c(3, 4, 1, 2, 7, 8, 5, 6))
g001 <- as.word(c(2, 1, 4, 3, 6, 5, 8, 7))

g100 * g010 == as.word(c(7, 8, 5, 6, 3, 4, 1, 2))
g100 * g001 == as.word(c(6, 5, 8, 7, 2, 1, 4, 3))
g001 * g010 == as.word(c(4, 3, 2, 1, 8, 7, 6, 5))
g100 * g010 * g001 == as.word(c(8, 7, 6, 5, 4, 3, 2, 1))

```

The list of the action element is
```{r}
g100
g010
g001
g100 * g010
g100 * g001
g001 * g010
g100 * g010 * g001

```

In the table above we proceed with the replacement: 

1 -> e
2 -> 001
3 -> 010
4 -> 011
5 -> 100
6 -> 101
7 -> 110
8 -> 111

|Beat  | e   | 001 | 010 | 011 | 100 | 101 | 110 | 111 |
|------|-----|-----|-----|-----|-----|-----|-----|-----|
|e     | e   | 001 | 010 | 011 | 100 | 101 | 110 | 111 |
|100   | 100 | 101 | 110 | 111 | e   | 001 | 010 | 011 |
|010   | 010 | 011 | e   | 001 | 110 | 111 | 100 | 101 |
|001   | 001 | e   | 011 | 010 | 101 | 100 | 111 | 110 |
|110   | 110 | 111 | 100 | 101 | 010 | 011 | e   | 001 |
|101   | 101 | 100 | 111 | 110 | 001 | e   | 011 | 010 |
|011   | 011 | 010 | 001 | e   | 111 | 110 | 101 | 100 |
|111   | 111 | 110 | 101 | 100 | 011 | 010 | 001 | e   |

Number of elements of S(8) that are composed of 4 2-cycles:
$8*7*6*5*4*3*2*1 \over 2^4 \times 4*3*2 = 7*3*5 = 105$

The recording space has 2^8 = 256 elements

A beat is the orbit of an element, it has a size of 1, 2, 4 or 8

    * If the beat size is 1
    
  Let f a recording of beat size 1. We have $\forall g \in C(2)^3 g \bullet f = f$
  We deduce that $\forall i \leq 8, f(i) = f(1)$
  So f is constant.
  
  So there are 2 beats of size 1
  
  1 1 1 1 1 1 1 1     
  and 
  0 0 0 0 0 0 0 0
  
  
    * If the beat size is 2
    
  