---
title: "C3"
author: "Manu"
date: "28 February 2017"
output: html_document
---

```{r setup, include=FALSE}
library(magrittr)
library(permutations)
source('test.R')
knitr::opts_chunk$set(echo = TRUE)
```

## C3 definition

We define C3 as a subgroup of R.

```{r, echo = FALSE}
rot_up <- cyc(c(1, 5, 7, 3), c(2, 6, 8, 4)) %>% as.cycle
rot_right <- cyc(c(1, 2, 6, 5), c(3, 4, 8, 7)) %>% as.cycle
rot_cck <- cyc(c(1, 2, 4, 3), c(5, 6, 8, 7)) %>% as.cycle
u <- rot_up
r <- rot_right
c <- rot_cck# elements
rr <- rot_right * rot_right
uu <- rot_up * rot_up
cc <- rot_cck * rot_cck
# all the group elements
SQ2 <- c(as.word(rr/rr), rr, uu, cc)
C3 <- c(SQ2,
  r*u*SQ2,
  (r*u)^2*SQ2)
C3
```

So C3 is a group with 12 elements. Including:

```{r, echo = FALSE}
cat('3 elements of order 2: \r\n')
C3[c(2, 3, 4)]
cat('8 elements of order 3: \r\n')
C3[5:12]
```

Since $C3 \subset R$ and SQ2 normal in R we deduce that SQ2 also normal in C3.

It has 3 classes: [e], [ru], [ru]^2 so it is isomorphic to Z/3Z. And $1 \mapsto ru$
defines an injection $Z/3Z \hookrightarrow C3$. So $C3 \sim Z/3Z \rtimes (Z/2Z)^2$.
We note Vx45 the subgroup {e, ru, (ru)^2}.

The action of Z/3Z on SQ2 can be defined by the action of ru on uu and rr, since ru generates Vx45, and uu and rr generates SQ2.

```{r}
all(uu^(r*u) == rr, rr^(r*u) == cc)
```

