---
title: "C3"
author: "Manu"
date: "28 February 2017"
output: html_document
---

```{r setup, include=FALSE}
library(magrittr)
library(permutations)
source('test.R')
knitr::opts_chunk$set(echo = TRUE)
```

## C3 definition

We define C3 as a subgroup of R.

```{r, echo = FALSE}
rot_up <- cyc(c(1, 5, 7, 3), c(2, 6, 8, 4)) %>% as.cycle
rot_right <- cyc(c(1, 2, 6, 5), c(3, 4, 8, 7)) %>% as.cycle
rot_cck <- cyc(c(1, 2, 4, 3), c(5, 6, 8, 7)) %>% as.cycle
u <- rot_up
r <- rot_right
c <- rot_cck# elements
rr <- rot_right * rot_right
uu <- rot_up * rot_up
cc <- rot_cck * rot_cck
ru <- r*u
# all the group elements
SQ2 <- c(as.word(rr/rr), rr, uu, cc)
C3 <- c(SQ2,
  r*u*SQ2,
  (r*u)^2*SQ2)
C3
```

So C3 is a group with 12 elements. Including:

```{r, echo = FALSE}
cat('3 elements of order 2: \r\n')
C3[c(2, 3, 4)]
cat('8 elements of order 3: \r\n')
C3[5:12]
```

Since $C3 \subset R$ and SQ2 normal in R we deduce that SQ2 also normal in C3.

It has 3 classes: [e], [ru], [ru]^2 so it is isomorphic to Z/3Z. And $1 \mapsto ru$
defines an injection $Z/3Z \hookrightarrow C3$. So $C3 \sim Z/3Z \rtimes (Z/2Z)^2$.
We note Vx45 the subgroup {e, ru, (ru)^2}.

The action of Z/3Z on SQ2 can be defined by the action of ru on uu and rr, since ru generates Vx45, and uu and rr generates SQ2.

```{r}
all(uu^(r*u) == rr, rr^(r*u) == cc)
```

## Other Semi-directs products $Z/3Z \rtimes (Z/2Z)^2$

We know that $\mathrm{Aut((Z/2Z)^2)} \sim S3 \sim Z/2Z \rtimes Z/3Z$

### Morphisms $Z/3Z \rightarrow S3$

Let $\phi: Z/3Z \rightarrow S3$. 

1. If $\phi$ is not injective then its kernel must be a normal subgroup of Z/3Z: $ker(\phi) = Z/3Z$ and $\forall a \in Z/3Z, \phi(a) = e_{S3}$

2. If $\phi is injective$
$\phi(1)$ is mapped to an element of order 3 in S3: either p or p2. We can verify that both give valid morphisms.

**Conlusion** We have 3 morphisms given respectively by $\phi(1) = e_{S3}, \phi(1) = p, \phi(1) = p^2$

### The semi-direct products

1. $\phi = e_{S3}$ and it gives the direct product $Z/3Z \times Z/2Z \times Z/2Z$
2. $\phi(1) = p$ gives the morphism $\phi(1)(1,0) = (0, 1)$ and $\phi(1)(0, 1) = (1, 1)$
And the semi-direct product is given by $p^iu^kv^n \cdot p^{i'}u^{k'}v^{n'} = p^{i+i'} \cdot u^nv^{k+n} \cdot u^{k'}v^{n'} = p^{i+i'} \cdot u^{n+k'}v^{k+n+n'}$
3. $\phi(1) = p2$ gives the morphism $\phi(1)(1, 0) = (1, 1)$ and $\phi(0, 1) = (1, 0)$
