---
title: "C6"
author: "Manu"
date: "January 18, 2018"
output: html_document
---

```{r setup, include=FALSE}
library(permutations)
library(DiagrammeR)
source('utils.R')
knitr::opts_chunk$set(echo = TRUE)
```

## Group defintion

Table of the group:

| C(6) | 0 | 1 | 2 | 3 | 4 | 5 |
|------|---|---|---|---|---|---|
| 0    | 0 | 1 | 2 | 3 | 4 | 5 |
| 1    | 1 | 2 | 3 | 4 | 5 | 0 |
| 2    | 2 | 3 | 4 | 5 | 0 | 1 |
| 3    | 3 | 4 | 5 | 0 | 1 | 2 |
| 4    | 4 | 5 | 0 | 1 | 2 | 3 |
| 5    | 5 | 0 | 1 | 2 | 3 | 4 |

Normal invariant series
$$C(1) \lhd C(3) \lhd C(6) \\
C(1) \lhd C(2) \lhd C(6)$$

With one C(2) and one C(3) composition factors

## Subgroups Lattice Graph

Since the group is abelian, the normal subgroup lattice is the same as the subgroup lattice.

```{r subgroup-lattice, echo = FALSE}
grViz('
graph cycle {
 rankdir="BT"

 node [shape = circle, style = filled, color = grey]

 0 [label = "0"]; C6 [label = "C(6)"]; 2 [label = "0, 2, 4"];
  3 [label = "0, 3"];

  0 -- 2 -- C6
  0 -- 3 -- C6
}
')
```




## Cycle Graph

```{r cycle-graph, echo = FALSE}
grViz('
graph cycle {
 graph[layout = neato]
 node [shape = circle, style = filled, color = grey, label=""]

 0 [color = red]; 1; 2; 3; 4; 5

  0 -- 1 -- 2 -- 3 -- 4 -- 5 -- 0
}
')
```


## Beat Action

There are 64 recordings.

Action associated with $C(1) \lhd C(3) \lhd C(6)$

The kernel is C(3) = {0, 2, 4}

|Beat  | 1 | 2 | 3 | 4 | 5 | 6 |
|------|---|---|---|---|---|---|
|e     | 1 | 2 | 3 | 4 | 5 | 6 |
|2     | 2 | 3 | 1 | 5 | 6 | 4 |
|4     | 3 | 1 | 2 | 6 | 4 | 5 |
|1     | 4 | 5 | 6 | 1 | 2 | 3 |
|3     | 5 | 6 | 4 | 2 | 3 | 1 |
|5     | 6 | 4 | 5 | 3 | 1 | 2 |

This is indeed an action because

```{r}
g_1 <- as.word(c(4, 5, 6, 1, 2, 3))
g_2 <- as.word(c(2, 3, 1, 5, 6, 4))
g_2 * g_1 == as.word(c(5,6,4,2,3,1))
g_2 * g_2 == as.word(c(3,1,2,6,4,5))
g_2 * g_2 * g_1 == as.word(c(6,4,5,3, 1, 2))
```

| Beat   | Size | Friends                                   | Negative |
|--------|------|-------------------------------------------|----------|
| 000000 | 1    | 0000                                      | No       |
| 100000 | 6    | 100000 010000 001000 000100 000010 000001 | No       |
| 100100 | 3    | 100100 010010 001001                      | No       |
| 100010 | 6    | 100010 010001 001100 010100 001010 100001 | No       |
| 110000 | 6    | 110000 011000 101000 000110 000011 000101 | No       |
| 110100 | 6    | 110100 011010 101001 100110 010011 001101 | No       |
| 111000 | 2    | 111000 000111                             | Itself   |
| 101010 | 6    | 101010 110001 011100 010101 001110 100011 | Itself   |
| 110010 | 6    | 110010 011001 101100 010110 001011 100101 | 110100   |