---
title: "Quaternion group"
author: "Manu"
date: "March 17, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(DiagrammeR)
```

## Quaternion group definition

We define the quaternion group QG by the following presentation on 2 generators and one relation.
$$<r^4, s^4, s^2 = r^2, s^3rs = r^3>$$

We notice that 

1. \<r\> is normal in QG
2. QG/\<r\> is generated by s and has order 2
3. \<r\> has order 4

So QG has order 8.

Here is the full multiplication table of QG

|     | e   | r   | r2  | r3  | s   | sr  | sr2 | sr3 |
|-----|-----|-----|-----|-----|-----|-----|-----|-----|
| e   | e   | r   | r2  | r3  | s   | sr  | sr2 | sr3 |
| r   | r   | r2  | r3  | e   | sr3 | s   | sr  | sr2 |
| r2  | r2  | r3  | e   | r   | sr2 | sr3 | s   | sr  |
| r3  | r3  | e   | r   | r2  | sr  | sr2 | sr3 | s   |
| s   | s   | sr  | sr2 | sr3 | r2  | r3  | e   | r   |
| sr  | sr  | sr2 | sr3 | s   | r   | r2  | r3  | e   |
| sr2 | sr2 | sr3 | s   | sr  | e   | r   | r2  | r3  |
| sr3 | sr3 | s   | sr  | sr2 | r3  | e   | r   | r2  |

Order of the elements:

| QG  | Order |
|-----|-------|
| e   | 1     |
| r   | 4     |
| r2  | 2     |
| r3  | 4     |
| s   | 4     |
| sr  | 4     |
| sr2 | 4     |
| sr3 | 4     |


## Normal sequence

\<r\> is maximally normal in QG and \<r\> ~ C(4)
So we get: 
$$C(2) \lhd C(4) \lhd QG$$

With 3 C(2) composition factors

## Cycle graph of QG

r is of order 4
s is of order 4
sr is of order 4
sr2 is of order 4
sr3 is of order 4

```{r echo = FALSE}
grViz('
graph cycle {
 graph[layout = neato]
 node [shape = circle, style = filled, color = grey, label=""]

 e [color = red]; r; r2 [label = "r2"]; r3;

  e -- r  -- r2 -- r3  -- e
  edge [color = "green"]
  e -- s  -- r2 -- sr2 -- e
  edge [color = "blue"]
  e -- sr -- r2 -- sr3 -- e
}
')
```


## Subgroups of QG

* Subgroups of order 2
{e, r2} is the only subgroup of order 2

* \<r\>, \<sr\>, \<s\> are subgroups of order 4

* If H is a subgroup of order 4 then it is isomorphic to C(4)

Let H be a subgroup of QG isomorphic to C(2)xC(2). Then H contains 3 elements
of order 2. Which is impossible since QD has only one element of order 2.

* \<r\>, \<sr\>, \<s\> are the only subgroups of order 4

H is isomorphic to C(4). Then H is uniquely defined by an element of order 4.
There are 6 such possibilities: r, r3, sr, sr3, s, s3. We verify for each that
they belong to one of \<r\>, \<sr\> or \<s\>.

Table of subgroups of QG

| QG   | Order |
|------|-------|
| {e}  | 1     |
| <r2> | 2     |
| <r>  | 4     |
| <sr> | 4     |
| <s>  | 4     |

Lattice graph of subgroups
```{r echo = FALSE}
grViz('
graph cycle {
 graph[layout = dot]
 node [shape = circle, style = filled, color = grey]

 e [color = red, label = "{e}"];
 r2 [label = "<r2>"];
 r [label = "<r>"];  sr [label = "<sr>"]; s [label = "<s>"];
 QG

 e -- r2
 r2 -- r
 r2 -- sr
 r2 -- s
 r -- QG
 sr -- QG
 s -- QG
}
')
```

## Normal subgroups of QG

\<r\> is normal in QG

We have $rsr^3 = sr^3r^3 = sr^2 = s^3 \in \langle s \rangle$. So \<s\> is normal in QG.

We also have 

  1. $s^3srs = rs = sr^3 = (sr)^3 \in \langle sr \rangle$ 

  2. $rsrr^3 = rs = sr^3 = (sr)^3 \in \langle sr \rangle$

So \<sr\> is also normal in QG. 

$\langle r^2 \rangle \subset \langle r \rangle$ so it \<r2\> is also normal in QG.

Conclusion all the subgroups of QG are normal. In particular its normal lattice iss the same as the subgroup lattice.
