---
output:
  html_document:
    toc: true
  pdf_document: default
---
:---
title: "S3"
author: "Manu"
date: "20 February 2017"
output: html_document
---

```{r setup, include=FALSE}
library(permutations)
library(igraph)
knitr::opts_chunk$set(echo = TRUE)
```

## S3 Group

The S3 group is the group of permutation on 3 elements. It has 6 elements. It is the group of all the automorphism on the set {1, 2, 3}.

Here is the full list of the elements of S3:

```{r elements, echo = FALSE}
e <- as.cycle()

a <- as.cycle(c(1,2))
b <- as.cycle(c(1,3))
c <- as.cycle(c(2,3))

p <- as.cycle(c(1,2,3))
p2 <- as.cycle(c(1,3,2))

cat(
  'e: ', as.character(e), '\r\n',
  'a: ', as.character(a), '\r\n',
  'b: ', as.character(b), '\r\n',
  'c: ', as.character(c), '\r\n',
  'p: ', as.character(p), '\r\n',
  'p^2: ', as.character(p2), '\r\n', sep = ''
  )
```

Table of S3:

| S3 | e  | p  | p2 | a  | b  | c  |
|----|----|----|----|----|----|----|
| e  | e  | p  | p2 | a  | b  | c  |
| p  | p  | p2 | e  | c  | a  | b  |
| p2 | p2 | e  | p  | b  | c  | a  |
| a  | a  | b  | c  | e  | p  | p2 |
| b  | b  | c  | a  | p2 | e  | p  |
| c  | c  | a  | b  | p  | p2 | e  |


The group \<p\> = {e, p, p2} is normal in S3:

```{r normal}
all(p^a == p2, p^b == p2, p^c == p2)
```

The classes of \<p\> in S3 are:

{e, p, p2}, {a, b, c}

The action of a on \<p\>: (p p2)

More detailed action of the group on \<p\>:

|S3|{e, p, p2}| {a, b, c}|
|--|----------|----------|
|e |{e, p, p2}|{a, b, c} |
|a |{a, b, c} |{e, p, p2}|
|b |{b, c, a} |{p2, e, p}|
|c |{c, a, b} |{p, p2, e}|
|p |{p, p2, e}|{c, a, b} |
|p2|{p2, e, p}|{b, c, a} |

## Subgroups of S3

\<p\> is isomorphic to Z/3Z and it has no non-trivial subgroups.

If G is a subgroup of S3 containing \<p\> then G*\<p\> is a subgroup of S3/\<p\>. S3/\<p\> is a group of order 2 so it is isomorphic to Z/2Z and it has no non trivial subgroup. The only subgroups containing p are {e} and S3.

If G is a sub-group not including \<p\>. Then inter(G, \<p\>) is {e}. because inter(G, \<p\>) is a subgroup of \<p\> and it cannot be \<p\> because G does no include \<p\>.

So G is included in {e, a, b, c}. G cannot contain more than 2 elements because any of products ab, bc, ... will yield an element of \<p\>.
So G is either {e, a}, {e, b}, {e, c}.

Conclusion: Here is the full list of the subgroups of S3:
{e}, {e, a}, {e, b}, {e, c}, {e, p, p2}, {e, a, b, c, p, p2}

**Lattice of all the subgroups**

```{r all-lattice, echo = FALSE}
g <- make_empty_graph() +
  vertex('S3') + 
  vertex('{e}') +
  vertex('<p>') + 
  vertex('<a>') +
  vertex('<b>') +
  vertex('<c>') +
  path('{e}', '<p>', 'S3') +
  path('{e}', '<a>', 'S3') +
  path('{e}', '<b>', 'S3') +
  path('{e}', '<c>', 'S3')

plot(g)
```

**Normal subgroups**

```{r normal-lattice, echo = FALSE}
g <- make_empty_graph() + 
vertex('S3') + 
vertex('{e}') +
vertex('<p>') + 
path('{e}', '<p>', 'S3')

plot(g)
```


## Actions of the group

### The natural action of S3 on {1, 2, 3}
This is a faithful action.

```{r, echo = FALSE}
a_f <- as.word(a, n = 3) %>% as.function
b_f <- as.word(b, n = 3) %>% as.function
c_f <- as.word(c, n = 3) %>% as.function
p_f <- as.function(p)

```

We can compute the orbits:

|S3| Orbits    |
|--|-----------|
|e |{1} {2} {3}|
|a |{1, 2} {3} |
|b |{1, 3} {2} |
|c |{1} {2, 3} |
|p |{1, 2, 3}  |
|p2|{1, 2, 3}  |

We can also compute the stabilizers:

|{1, 2, 3}|Stabilizer|
|---------|----------|
|1        |\<c\>       |
|2        |\<b\>       |
|3        |\<a\>       |

### Action on the 'edges'

We consider the action of S3 on this set: {{1, 2}, {1, 3}, {2, 3}}

* Action of \<p\>

```{r p-action, echo=FALSE}
all(
  p_f(c(1, 2)) == c(2, 3),
  p_f(c(1,3)) == c(2, 1),
  p_f(c(2, 3)) == c(3, 1))
```

* Action of a:
```{r}
all(
  a_f(c(1, 2)) == c(2, 1),
  a_f(c(1,3)) == c(2, 3),
  a_f(c(2, 3)) == c(1, 3))
```

Since a and p generate S3 we deduce that the action is *faithfull*.

Orbits

|S3| Orbits          |
|--|-----------------|
|e |{1_2} {1_3} {2_3}|
|a |{1_2} {1_3, 2_3} |
|b |{1_2, 2_3} {1_3} |
|c |{1_2, 1_3} {2, 3}|
|p |{1_2, 2_3, 1_3}  |
|p2|{1_2, 2_3, 1_3}  |

Stabilizers

### Other actions


* The action of S3 on S3/\<p\> is not faithfull.
S3/\<p\> = {{e, p, p2}, {a, b, c}}
Any element of \<p\> is sent to the identity.
Any other element is sent to ({e, p, p2} {a, b, c})

* There is also an action of S3 on {a, b, c}.

* And an action of S3 on \<p\>

### All the transitive, faithfull actions of S3

Let $\phi$ be a transitive, faithfull set action $S3 \rightarrow S(\mathcal{A})$. Where $\mathcal{A}$ is any set.
We note: $g \bullet x = \phi(g)(x)$ for $g \in S3, a \in \mathcal{A}$

First we would like to show that $\mathcal{A}$ has 6 elements.

Let $a_0 \in \mathcal{A}$, we let $f_{a_0}: S3 \rightarrow \mathcal{A}$ defined by $f_{a_0}(g) = g \bullet a_0$

Since the action is transitive $f_{a_0}$ is surjective.
Let's $g_1, g_2 \in \mathrm{S3}$. 
$$f_{a_0}(g_1) = f_{a_0}(g_2) \iff g_1 \bullet a_0 = g_2 \bullet a_0 \iff g_1 g_2^{-1} \bullet a_0 = a_0$$


### How to see if 2 actions are isomorphic?
In all this file G is a finite group and S is a set.

An action is a group morphism $\phi_1: G \rightarrow Aut(S)$ where S is some set.
Let $\phi_2: G \rightarrow Aut(S2)$ be another group action.
2 action are isomorphic iff there exist a set isomorphism $f: S \rightarrow S2$
such that $$\forall g \in G, f(\phi_1(g)(x)) = \phi_2(g)(f(x))$$

The next property shows that if we have verified the property on f for $g_1, g_2 \in G$, then the property holds for $g_1g_2$.
$$f(\phi_1(g_1g_2)(x)) = f(\phi_1(g_1)(\phi_1(g_2)(x)))
= \phi_2(g_1)(f(\phi_1(g_2)x)) = \phi_2(g_1)(\phi_2(g_2)(fx))
= \phi_2(g_1g_2)(fx)$$

That means that we only need to verify the property on a set of generators for G.


### Example
Let's show that the action on {1, 2, 3} is isomorphic to the action on {{1, 2}, {1, 3}, {2, 3}} by taking 
f: 1 -> {2, 3}
f: 2 -> {1, 3}
f: 3 -> {1, 2}

f(p.1) = f(2) = {1, 3}
p.f(1) = p.{2, 3} = {1, 3}

f(p.2) = f(3) = {1, 2}
p.f(2) = p.{1, 3} = {1, 2}

f(a.1) = f(2) = {1, 3}
a.f(1) = a.{2, 3} = {1, 3}

f(a.3) = f(3) = a.f(3)
=> f(3) = {1, 2}

f(b.2) = f(2) = b.f(2)
=> f(2) = {1, 3}

f(c.1) = f(1) = c.f(1)
=> f(1) = {2, 3}

So f is an isomorphism of actions.


# Automorphism group of S3

## Inner automorphisms

Conjugation by an element creates an automorphism.
The conjugation by p has the following effect:
```{r conjugation, echo = FALSE}
cat(
  'a^p: ', a^p %>% as.cycle %>% as.character, '\r\n',
  'b^p: ', b^p %>% as.cycle %>% as.character, '\r\n',
  'c^p: ', c^p %>% as.cycle %>% as.character, '\r\n',
  'p^p: ', p^p %>% as.cycle %>% as.character, '\r\n',
  'p2^p: ', (p*p)^p %>% as.cycle %>% as.character, '\r\n', sep = ''
)
```

So conj(p) = (a c b)

Here is the conjugation by p2:

```{r conjugation-p2, echo = FALSE}
cat(
  'a^p2: ', a^(p*p) %>% as.cycle %>% as.character, '\r\n',
  'b^p2: ', b^(p*p) %>% as.cycle %>% as.character, '\r\n',
  'c^p2: ', c^(p*p) %>% as.cycle %>% as.character, '\r\n',
  'p^p2: ', p^(p*p) %>% as.cycle %>% as.character, '\r\n',
  'p2^p2: ', (p*p)^(p*p) %>% as.cycle %>% as.character, '\r\n', sep = ''
)
```

So conj(p2) = (a b c)

Here is the conjugation by a:
```{r conjugation-a, echo = FALSE}
cat(
  'a^a: ', a^a %>% as.cycle %>% as.character, '\r\n',
  'b^a: ', b^a %>% as.cycle %>% as.character, '\r\n',
  'c^a: ', c^a %>% as.cycle %>% as.character, '\r\n',
  'p^a: ', p^a %>% as.cycle %>% as.character, '\r\n',
  'p2^a: ', (p*p)^a %>% as.cycle %>% as.character, '\r\n', sep = ''
)
```

So conj(a) = (b c) (p p2)

Here is the conjugation by b:

```{r conjugation-b, echo = FALSE}
cat(
  'a^b: ', a^b %>% as.cycle %>% as.character, '\r\n',
  'b^b: ', b^b %>% as.cycle %>% as.character, '\r\n',
  'c^b: ', c^b %>% as.cycle %>% as.character, '\r\n',
  'p^b: ', p^b %>% as.cycle %>% as.character, '\r\n',
  'p2^b: ', (p*p)^b %>% as.cycle %>% as.character, '\r\n', sep = ''
)
```

So conj(b) = (a c) (p p2)

Here is the conjugation by c:
```{r conjugation-c, echo = FALSE}
cat(
  'a^c: ', a^c %>% as.cycle %>% as.character, '\r\n',
  'b^c: ', b^c %>% as.cycle %>% as.character, '\r\n',
  'c^c: ', c^c %>% as.cycle %>% as.character, '\r\n',
  'p^c: ', p^c %>% as.cycle %>% as.character, '\r\n',
  'p2^c: ', (p*p)^c %>% as.cycle %>% as.character, '\r\n', sep = ''
)
```

So conj(c) = (a b)(p p2)

So together with the identity there are 6 inner automorphisms.

### Structure of the automorphism group

We will show that Inner(S3) = S3
conj(a)^2 = e
conj(p)^3 = e
conj(p)^conj(a) = conj a-1 o conj p o conj a = conj a-1pa = conj p2 = conj(p)^2

Since the generators of Inner(S3) have the same relations as those of S3 we know the 2 groups are isomorphic.

## Total group or isomorphism

We can show that there are no outer isomorphism. Indeed an automorphism maps elements to elements of the same order. There are thus 2 possibilities for mapping p and 3 possibilities for mapping a. Since a and p generate S3 this 2 images will define uniquely a mapping. There are at most 3*2 automorphisms and according to the previous paragraphs we already have 6 inner automorphisms. So all the automorphism are inner.

## Matrix representation of an automorphism

Since an automorphism is uniquely defined by its image of the generators a and p we might be able to write them as a matrix. Indeed the reason we are able to write linear operators as matrix is because they are defined by their action on a basis which is a generator set of the whole vector space.

First lets try to write elements of S3 as vectors:
we note a = (1, 0) and p = (0, 1)
(0, x) \* (1, m)  = p^x \* a \* p^m = a \* a p^x \* a p^m = (1, 2*x+m)

(g1, n1) \* (g2, n2) = g1n1g2n2 = g1g2inv(g2)n1g2n2 = g1g2n1^g2n2 
= (g1g2, n1^g2\*n2)

inv(g1) \* (g2, n2) \* g1 = inv(g1) \* (g2*g1, n2^g1) = (g2^g1, n2^g1)
conj(g1) = 
g1      e
e       g1
