---
title: "Symmetric Groups"
author: "Manu"
date: "June 29, 2017"
output: 
  html_document:
    toc: true
---

```{r setup, include=FALSE}
library(DiagrammeR)
knitr::opts_chunk$set(echo = TRUE)
```

## Definition

**Symmetric group**
The Symmetric Group on a finite set A is defined as the set of all invertible functions $A \rightarrow A$ together with the opposite composition as the operation. To sets with the same number of elements will give isomorphic groups. We note Sn the symmetric gorup on n elements.


**Cycle**
A cycle c in Sn is a permutation such that the action of <c> on A has only one orbit of size > 1.

**Functoriality**
Let **Setinq** be the category of all finite sets, with only injective functions as the morphisms. This is a category because the identity is injective and the composition of 2 injections is itself an injection.

Then Sym is a functor $\mathrm{Setinq} \rightarrow \mathrm{Group}$

Indeed if $f: A \hookrightarrow B$ then we can define $Sym(f): Sym(A) \hookrightarrow Sym(B)$ by:

$$\mathrm{Sym}(f)(g)(b) = \left \{ \begin{array}{ll} 
  b, b \notin Im(f) \\
  f(g(f^{-1}(b))), \mathrm{otherwise} 
  \end{array} \right .$$

## Examples

### S4

S4 has 24 elements

| Order        | Number | Elements |
|--------------|--------|----------|
| 1            |   1    |   id     |
| 2            |   9    | (1 2), (1 3), (1, 4), (2, 3), (2, 4), (3, 4), (1 2) (3 4), (1 3) (2 4), (1 4) (2 3) |
| 3            |   8    | (1 2 3), (1 3 2), (1 2 4), (1 4 2), (1 3 4), (1 4 3), (2 3 4), (2 4 3) |
| 4            |    6   | (1 2 3 4), (1 4 3 2), (1 2 4 3), (1 3 4 2), (1 3 2 4), (1 4 2 3) |

```{r}
grViz('
graph S4 {
  graph [layout = twopi, size="7.75,10.25", ranksep=3.0, nodesep=2.0]

  e
  p_1_2 [label = "(1 2)"]
  p_1_3 [label = "(1 3)"]
  p_1_4 [label = "(1 4)"]
  p_2_3 [label = "(2 3)"]
  p_2_4 [label = "(2 4)"]
  p_3_4 [label = "(3 4)"]
  
  p_1_2_3 [label = "(1 2 3)"]
  p_1_3_2 [label = "(1 3 2)"]
  p_1_2_4 [label = "(1 2 4)"]
  p_1_4_2 [label = "(1 4 2)"]
  p_1_3_4 [label = "(1 3 4)"]
  p_1_4_3 [label = "(1 4 3)"]
  p_2_3_4 [label = "(2 3 4)"]
  p_2_4_3 [label = "(2 4 3)"]

  p_1_2_3_4 [label = "(1 2 3 4)"]
  p_1_3_T_2_4 [label = "(1 3) (2 4)"]
  p_1_4_3_2 [label = "(1 4 3 2)"]

  p_1_2_4_3 [label = "(1 2 4 3)"]
  p_1_4_T_2_3 [label = "(1 4) (2 3)"]
  p_1_3_4_2 [label = "(1 3 4 2)"]

  p_1_3_2_4 [label = "(1 3 2 4)"]
  p_1_2_T_3_4 [label = "(1 2) (3 4)"]
  p_1_4_2_3 [label = "(1 4 2 3)"]


  e -- p_1_2
  e -- p_1_3
  e -- p_1_4
  e -- p_2_3
  e -- p_2_4
  e -- p_3_4

  e -- p_1_2_3
  e -- p_1_3_2
  e -- p_1_2_4
  e -- p_1_4_2
  e -- p_1_3_4
  e -- p_1_4_3
  e -- p_2_3_4
  e -- p_2_4_3

  p_1_2_3 -- p_1_3_2
  p_1_2_4 -- p_1_4_2
  p_1_3_4 -- p_1_4_3
  p_2_3_4 -- p_2_4_3

  e -- p_1_2_3_4
  e -- p_1_4_3_2
  e -- p_1_2_4_3
  e -- p_1_3_4_2
  e -- p_1_3_2_4
  e -- p_1_4_2_3

  p_1_2_3_4 -- p_1_3_T_2_4
  p_1_4_3_2 -- p_1_3_T_2_4

  p_1_2_4_3 -- p_1_4_T_2_3
  p_1_3_4_2 -- p_1_4_T_2_3

  p_1_3_2_4 -- p_1_2_T_3_4
  p_1_4_2_3 -- p_1_2_T_3_4

}
')
```


### S5

S5 has 120 elements. 
It has 5\*4/2 = 10 permutations. 
It has 5\*4\*3 / 3 = 20 3-cycles
It has 5\*4\*3\*2 / 4 = 30 4-cycles
4\*3*2 = 24 5-cycles

| Order        | 1 | 2 | 3  | 4   | 5   | 6    | Total     |
|--------------|---|---|----|-----|-----|------|-----------|
| # elements   | 1 | ? | 20 | 30  | 24  | ?    | 75        |

### S6

S6 has 720 elements

It has 6\*5/2 = 15 permutations. 
It has 6\*5\*4 / 3 = 40 3-cycles
It has 6\*5\*4\*3 / 4 = 90 4-cycles
6\*5\*4\*3\*2 / 5 = 144 5-cycles
6\*5\*4\*3\*2 / 6 = 120 6-cycles

## Properties

The symmetric group has n! elements.

**proof**
Let f be in Sn. There are n choices for f(1), n-1 choices for f(2) (because f(1) must be different to f(2)), n-3 choices for f(3)... all the way to 1. So there is a total of n(n-1)(n-2)...1 = n! choices for f.

| n        | 2 | 3 | 4  | 5   | 6   | 7    | 8     | 9       | 10        |
|----------|---|---|----|-----|-----|------|-------|---------|-----------|
| Card(Sn) | 2 | 6 | 24 | 120 | 720 | 5040 | 40320 | 362,880 | 3,628,800 |

Some elements of the symmetric group can have order >= n

**example**
In S5, (1 2 3)(4 5) has order 6.


## Signature

Let Cou be the set of all $(x, y) \in A^2 / x \neq y$. Then Cou is a C(2)-set together with the action $1+(x, y) = (y, x)$.

Let P be the set of all unordered pairs $\{a, b\}, a,b \in A^2, a \neq b$.

We call $\pi_P: \mathrm{Cou} \rightarrow P$ the natural projection on P that drops the order on the ordered pair.

For example, if A = {1, 2, 3} then Cou = {(1, 2), (1, 3), (2, 3), (2, 1), (3, 1), (3, 2)}

And P = {{1, 2}, {1, 3}, {2, 3}}

In general

| Size of A | Size Cou | Size P   |
|-----------|----------|----------|
| 3         | 6        | 3        |
| 4         | 12       | 6        |
| 5         | 20       | 10       |
| 6         | 30       | 15       |
| 7         | 42       | 21       |
| 8         | 56       | 28       |
| 9         | 72       | 36       |
| n         | n(n-1)   | n(n-1)/2 |



### Definition of the Signature

Let $\pi_2: Cou \rightarrow C(2)$ be a C(2)-map and let $s \in S(A)$ we write $\sigma(s) = \sum_{(x, y) \in {\pi_2}^{-1}(0)} \pi_2(s(x), s(y))$. This quantity does not depend on $\pi_2$

### Maps P -> Cou

Let I be the set of all maps $i: P \rightarrow Cou$ such that $i \circ_{op} \pi_P = id_P$

For all {x, y} in P, we have $\pi(i({x, y})) = \{x, y\}$ so $i(\{x, y\}) = \left [  \begin{array}{ll} (x, y) \\ or \\  (y, x) \end{array} \right .$

We easily verify that any map verrifying this property is in I. There are $n(n-1) \over 2$ elements in P and for each we have 2 possibilities. So I contains $2^{n(n-1) \over 2}$ elements.

For n=3 I contains 2^3 = 8 elements.

| I | {1, 2} | {1, 3} | {2, 3} |
|---|--------|--------|--------|
| 1 | (1, 2) | (1, 3) | (2, 3) |
| 2 | (1, 2) | (1, 3) | (3, 2) |
| 3 | (1, 2) | (3, 1) | (2, 3) |
| 4 | (1, 2) | (3, 1) | (3, 2) |
| 5 | (2, 1) | (1, 3) | (2, 3) |
| 6 | (2, 1) | (1, 3) | (3, 2) |
| 7 | (2, 1) | (3, 1) | (2, 3) |
| 8 | (2, 1) | (3, 1) | (3, 2) |

We can also define an action of S(A) unto I by
$$(\{x, y\})i \cdot g = \left \{ \begin{array}{ll}
(x, y), \mathrm{if} (x, y) \in g(i(P)) \\
(y, x), \mathrm{if} (y, x) \in g(i(P)) \end{array} \right .$$

We can easily see that only one of the 2 conditions can be true at the same time, and that one always has to be true at anytime.

We can also define $i \cdot g$ as the only application in I that has values in g(i(P)).

Indeed g(i(P)) is a set a representative for the equivalence relation $(x,y) \sim (y, x)$. Because:

1. For any element (x, y), we can find (a, b) in g(i(P)) such that g(a,b) = (x, y) or g(a, b) = (y, x)
2. For any class {(x, y), (y, x)} there exist only one g(a, b) in g(i(P)) such that g(a, b) is in this class.

**Example n=3**

Action of a of S3 on $i_1$:

| I                   | {1, 2} | {1, 3} | {2, 3} | Result  |
|---------------------|--------|--------|--------|---------|
| $i_1$               | (1, 2) | (1, 3) | (2, 3) |  $i_1$  |
| $i_1 \cdot (1 2)$   | (2, 1) | (1, 3) | (2, 3) |  $i_5$  |
| $i_1 \cdot (1 3)$   | (2, 1) | (3, 1) | (3, 2) |  $i_8$  |
| $i_1 \cdot (2 3)$   | (1, 2) | (1, 3) | (3, 2) |  $i_2$  |
| $i_1 \cdot (1 2 3)$ | (2, 1) | (3, 1) | (2, 3) |  $i_7$  |
| $i_1 \cdot (1 3 2)$ | (1, 2) | (3, 1) | (3, 2) |  $i_4$  |

$O(i_1) = \{i_1, i_2, i_4, i_5, i_7, i_8\}$
$\mathrm{Stab}(i_1) = \{e\}$

As we can se $i_3$ and $i_6$ are not there

| I                   | {1, 2} | {1, 3} | {2, 3} | Result  |
|---------------------|--------|--------|--------|---------|
| $i_3$               | (1, 2) | (3, 1) | (2, 3) |  $i_3$  |
| $i_3 \cdot (1 2)$   | (2, 1) | (1, 3) | (3, 2) |  $i_6$  |
| $i_3 \cdot (1 3)$   | (2, 1) | (1, 3) | (3, 2) |  $i_6$  |
| $i_3 \cdot (2 3)$   | (2, 1) | (1, 3) | (3, 2) |  $i_6$  |
| $i_3 \cdot (1 2 3)$ | (1, 2) | (3, 1) | (2, 3) |  $i_3$  |
| $i_3 \cdot (1 3 2)$ | (1, 2) | (3, 1) | (2, 3) |  $i_3$  |

$O(i_3) = \{i_3 i_6\}$
$\mathrm{Stab}(i_3) = \{e, (1 2 3), (1 3 2)\}$


| I                   | {1, 2} | {1, 3} | {2, 3} | Result  |
|---------------------|--------|--------|--------|---------|
| $i_6$               | (2, 1) | (1, 3) | (3, 2) |  $i_6$  |
| $i_6 \cdot (1 2)$   | (1, 2) | (3, 1) | (2, 3) |  $i_3$  |
| $i_6 \cdot (1 3)$   | (1, 2) | (3, 1) | (2, 3) |  $i_3$  |
| $i_6 \cdot (2 3)$   | (1, 2) | (3, 1) | (2, 3) |  $i_3$  |
| $i_6 \cdot (1 2 3)$ | (2, 1) | (1, 3) | (3, 2) |  $i_6$  |
| $i_6 \cdot (1 3 2)$ | (2, 1) | (1, 3) | (3, 2) |  $i_6$  |

8 = 6/1 + 6/3

**Example n=4**

For n=4 I contains 2^6 = 64 elements.

Each app i contains 6 pairs.
Each action involves 24 elements

Let's consider $O(i)$:

|             | {1, 2} | {1, 3} | {1, 4} | {2,3}  | {2, 4} | {3, 4} |
|-------------|--------|--------|--------|--------|--------|--------|
| i           | (1, 2) | (1, 3) | (4, 1) | (2, 3) | (2, 4) | (3, 4) | 
| i.(1 2)     | (2, 1) | (1, 3) | (1, 4) | (2, 3) | (4, 2) | (3, 4) |
| i.(1 2 3 4) | (1, 2) | (3, 1) | (4, 1) | (2, 3) | (2, 4) | (3, 4) |
...

Let's consider

|             | {1, 2} | {1, 3} | {1, 4} | {2,3}  | {2, 4} | {3, 4} |
|-------------|--------|--------|--------|--------|--------|--------|
| i           | (1, 2) | (3, 1) | (1, 4) | (2, 3) | (2, 4) | (3, 4) |
| i.(1 2 3)   | (1, 2) | (3, 1) | (1, 4) | (2, 3) | (2, 4) | (3, 4) | 

## Chains of couples

Let g be any element in S(A) and (x, y) any pair we can define a chain:

(x, y) -> (gx, gy) -> $(g^2 \cdot x, g^2 \cdot y)$ ->  ... -> (x, y)

**Examples**

Let A = {1, 2, 3} and g = (1 2 3)

We have the following chains:

(1, 2) -> (2, 3) -> (3, 1) -> (1, 2)
(2, 1) -> (3, 2) -> (1, 3) -> (2, 1)

We can see that up to a reordering of the chain, we only have 2 chains, and they are opposite by the action of C(2).

We can also see that up to the action of C(2) we have 3 chains
(1, 2) -> (2, 3) -> (3, 1) -> (1, 2)
(2, 3) -> (3, 1) -> (1, 2) -> (2, 3)
(3, 1) -> (1, 2) -> (2, 3) -> (3, 1)

If we quotient by the 2 actions we get only 1 chain:

(1, 2) -> (2, 3) -> (3, 1) -> (1, 2) 0


---

Now for g = (1 2)

(1, 2) -> (2, 1) -> (1, 2)
(1, 3) -> (2, 3) -> (1, 3)
(3, 1) -> (3, 2) -> (3, 1)

Here we have one chain symmetric to itself and a pair of opposite chains, up to reordering of the chain.
We can also see that up to the action of C(2) we have 3 chains

(1, 2) -> (2, 1) -> (1, 2)
(1, 3) -> (2, 3) -> (1, 3)
(2, 3) -> (1, 3) -> (2, 3)

Now if we quotient by the 2 actions, we get only 2 chains:

(1, 2) -> (2, 1) -> (1, 2) 1
(1, 3) -> (2, 3) -> (1, 3) 0

---

Let A = {1, 2, 3, 4} and g = (1 2 3 4)

(1, 2) -> (2, 3) -> (3, 4) -> (4, 1) -> (1, 2) = 0
(1, 3) -> (2, 4) -> (3, 1) -> (4, 2) -> (1, 3) = 1
(1, 4) -> (2, 1) -> (3, 2) -> (4, 3) -> (1, 4) = 0
(2, 3) -> (3, 4) -> (4, 1) -> (1, 2) -> (2, 3) = 0
(2, 4) -> (3, 1) -> (4, 2) -> (1, 3) -> (2, 4) = 1
(3, 4) -> (4, 1) -> (1, 2) -> (2, 3) -> (3, 4) = 0

Up to C(2)

If we also quotient by the action of g:

(1, 2) -> (2, 3) -> (3, 4) -> (4, 1) -> (1, 2) = 0
(1, 3) -> (2, 4) -> (3, 1) -> (4, 2) -> (1, 3) = 1

---

Let A = {1, 2, 3, 4} and g = (1 2) (3 4)

Quotient by the action of g:

(1, 2) -> (2, 1) -> (1, 2)

(1, 3) -> (2, 4) -> (1, 3)
(3, 1) -> (4, 2) -> (3, 1)

(1, 4) -> (2, 3) -> (1, 4)
(4, 1) -> (3, 2) -> (4, 1)

(3, 4) -> (4, 3) -> (3, 4)

Now we quotient again by the action of C(2).

(1, 2) -> (2, 1) -> (1, 2) 1
(1, 3) -> (2, 4) -> (1, 3) 0
(1, 4) -> (2, 3) -> (1, 4) 0
(3, 4) -> (4, 3) -> (3, 4) 1

---

Let A = {1, 2, 3, 4} and g = (1 2 3)

Quotient by the action of g

(1, 2) -> (2, 3) -> (3, 1) -> (1, 2)
(2, 1) -> (3, 2) -> (1, 3) -> (2, 1)

(1, 4) -> (2, 4) -> (3, 4) -> (1, 4)
(4, 1) -> (4, 2) -> (4, 3) -> (4, 1)

If we quotient again by the action of C(2) we get 2 chains:

(1, 2) -> (2, 3) -> (3, 1) -> (1, 2)
(1, 4) -> (2, 4) -> (3, 4) -> (1, 4)

---

Let A = {1, 2, 3, 4, 5} and g = (1 2 3 4 5)

If we don't quotient by any action we get 20 chains (too much)

If we quotient by the action of g we get 4 chains

(1, 2) -> (2, 3) -> (3, 4) -> (4, 5) -> (5, 1) -> (1, 2)
(2, 1) -> (3, 2) -> (4, 3) -> (5, 4) -> (1, 5) -> (2, 1)

(1, 3) -> (2, 4) -> (3, 5) -> (4, 1) -> (5, 2) -> (1, 3)
(3, 1) -> (4, 2) -> (5, 3) -> (1, 4) -> (2, 5) -> (3, 1)

If we quotient by C(2) we get 2 chains:

(1, 2) -> (2, 3) -> (3, 4) -> (4, 5) -> (5, 1) -> (1, 2) 0
(1, 3) -> (2, 4) -> (3, 5) -> (4, 1) -> (5, 2) -> (1, 3) 0

---

Let A = {1, 2, 3, 4, 5} and g = (1 2 3) (4 5)

We quotient by C(2) first, so we get 10 chains:

(1, 2) -> (2, 3) -> (3, 1) -> (1, 2)
(1, 3) -> (2, 1) -> (3, 2) -> (1, 3)
(1, 4) -> (2, 5) -> (3, 4) -> (1, 5) -> (2, 4) -> (3, 5) -> (1, 4)
(1, 5) -> (2, 4) -> (3, 5) -> (1, 4) -> (2, 5) -> (3, 4) -> (1, 5)
(2, 3) -> (3, 1) -> (1, 2) -> (2, 3)
(2, 4) -> (3, 5) -> (1, 4) -> (2, 5) -> (3, 4) -> (1, 5) -> (2, 4)
(2, 5) -> (3, 4) -> (1, 5) -> (2, 4) -> (3, 5) -> (1, 4) -> (2, 5)
(3, 4) -> (1, 5) -> (2, 4) -> (3, 5) -> (1, 4) -> (2, 5) -> (3, 4)
(3, 5) -> (1, 4) -> (2, 5) -> (3, 4) -> (1, 5) -> (2, 4) -> (3, 5)
(4, 5) -> (5, 4) -> (4, 5)

Now we quotient by g we get 3 chains:

(1, 2) -> (2, 3) -> (3, 1) -> (1, 2) 0
(1, 4) -> (2, 5) -> (3, 4) -> (1, 5) -> (2, 4) -> (3, 5) -> (1, 4) 0
(4, 5) -> (5, 4) -> (4, 5) 1

---

Let A = {1, 2, 3, 4, 5} and g = (1 2 3) 

Quotient by C(2) first:
(1, 2) -> (2, 3) -> (3, 1) -> (1, 2)
(1, 3) -> (2, 1) -> (3, 2) -> (1, 3)
(1, 4) -> (2, 4) -> (3, 4) -> (1, 4)
(1, 5) -> (2, 5) -> (3, 5) -> (1, 5)
(2, 3) -> (3, 1) -> (1, 2) -> (2, 3)
(2, 4) -> (3, 4) -> (1, 4) -> (2, 4)
(2, 5) -> (3, 5) -> (1, 5) -> (2, 5)
(3, 4) -> (1, 4) -> (2, 4) -> (3, 4)
(3, 5) -> (1, 5) -> (2, 5) -> (3, 5)
(4, 5) -> (4, 5)

Quotient by <g>:
(1, 2) -> (2, 3) -> (3, 1) -> (1, 2) 0
(1, 4) -> (2, 4) -> (3, 4) -> (1, 4) 0
(1, 5) -> (2, 5) -> (3, 5) -> (1, 5) 0
(4, 5) -> (4, 5) 0


---

Let A = {1, 2, 3, 4, 5} and g = (4 5)

Quotient by C(2):

(1, 2) -> (1, 2) 
(1, 3) -> (1, 3) 
(1, 4) -> (1, 5) -> (1, 4) 
(1, 5) -> (1, 4) -> (1, 5) 
(2, 3) -> (2, 3) 
(2, 4) -> (2, 5) -> (2, 4) 
(2, 5) -> (2, 4) -> (2, 5) 
(3, 4) -> (3, 5) -> (3, 4) 
(3, 5) -> (3, 4) -> (3, 5) 
(4, 5) -> (5, 4) -> (4, 5) 

Quotient by <g>:

(1, 2) -> (1, 2) 
(1, 3) -> (1, 3) 
(1, 4) -> (1, 5) -> (1, 4) 
(2, 3) -> (2, 3) 
(2, 4) -> (2, 5) -> (2, 4) 
(3, 4) -> (3, 5) -> (3, 4) 
(4, 5) -> (5, 4) -> (4, 5) 

They all have signature 0 except the last one.

### 3 Chains

### Definition of 2-chains

Let A be any finite set, g any element in S(A). Let $(x, y) \in Cou(A)$ we can define **chain((x,y), g)** as the serie $(g^n(x), g^n(y))_{n \in N}$

**Properties**

1. $\forall n \in N, (g^n(x), g^n(y)) \in Cou(A)$
2. La serie est periodique: there exists n0 such that $(g^n(x), g^n(y)) = (g^{n_0+n}(x), g^{n_0 +n}(y))$
3. A g-chain is uniquely defined by its first value, so we note it $(x, y)_g$
**Definition**

We define **Chain(g)** has the set of all chains {chain((x,y), g) / (x, y) in Cou}

|Chain(g)| = n(n-1) where n = |A|

We have one **action of C(2)** on Chain(g) defined by $1+(g^n(x), g^n(y))_{n \in N} = (g^n(y), g^n(x))_{n \in N}$. We easily verify that both are g-chains.

We also have one **action of <g>** on Chain(g) defined by $1+(g^n(x), g^n(y))_{n \in N} = (g^n(g(x)), g^n(g(y)))_{n \in N}$

Both actions commute: 1+(g.c) = g.(1+c)
Because of that we can define a g action on the orbits by C(2).

Let's call the reduced chains the following set:
$RChain(g) = O(O(Chain(g), C(2)), <g>)$

2 chains belong to the same reduced chain iff:
  $(x, y)_g \sim (x', y')_g \iff \exists a \in C(2), h \in <g> / (x, y) = a+g\cdot (x', y')$
  
Let $(x, y)_g$ be a g-chain. We define a function $\sigma: Cou(A) \rightarrow C(2)$ by the following. $\sigma((x, y)_g) = 1 \iff \exists h \in <g> / h \cdot (x, y) = (y, x)$

We have the following properties: $\sigma(1+(x,y)_g) = \sigma((x,y)_g)$ and $\sigma(g(x,y)_g) = \sigma((x,y)_g)$ 
The first is obvious. For the second one, let h such that (hx, hy) = (y, x)
then (hgx, hgy) = (gx, gy) because h and g commute.

So $\sigma$ is well defined on RChain(g) and we can write $\sigma(g) = \sum_{c \in Rchain}\sigma(c)$

Let $g_1, g_2 \in S(A), (x, y) \in Cou$
Suppose that there exists h such that $(hg1x, hg1y) \cdot g_2 = (g1y, g1x)$ then $(x, y) \cdot{g_1g_2} = (g1y, g1x) \cdot g_2 = (hg1y, hg1x)$

## Functions $G \times G \rightarrow C(2)$

Let f, g in Hom(G2, C(2)) then we can define (f+g)(x, y) = f(x, y) + g(x, y)

For $(x_0, y_0) \in P$ we define $\mathcal{I}_{x_0, y_0}$ in Hom(G2, C(2)) to be the function equal to 1 on $(x_0, y_0)$ and 0 everywhere else.

