---
title: "Z/4Z"
author: "Manu"
date: "3 March 2017"
output: html_document
---

```{r setup, include=FALSE}
library(permutations)
library(DiagrammeR)
source('utils.R')
knitr::opts_chunk$set(echo = TRUE)
```

## Group defintion

Table of the group:

| Z/4Z | 0 | 1 | 2 | 3 |
|------|---|---|---|---|
| 0    | 0 | 1 | 2 | 3 |
| 1    | 1 | 2 | 3 | 0 |
| 2    | 2 | 3 | 0 | 1 |
| 3    | 3 | 0 | 1 | 2 |

Putting the subgroup {0, 2} first:

| Z/4Z | 0 | 2 | 1 | 3 |
|------|---|---|---|---|
| 0    | 0 | 2 | 1 | 3 |
| 2    | 2 | 0 | 3 | 1 |
| 1    | 1 | 3 | 2 | 0 |
| 3    | 3 | 1 | 0 | 2 |


$C(1) \lhd C(2) \lhd C(4)$ With two C(2) composition factors
This decomposition is unique


## Subgroups

```{r subgroup-lattice, echo = FALSE}
grViz('
graph cycle {
 graph[layout = neato]
 node [shape = circle, style = filled, color = grey]

 0 [label = "0"]; C4 [label = "C(4)"]; 2 [label = "{0, 2}"];

  0 -- 2 -- C4
}
')
```



## Graphs

```{r cycle-graph, echo = FALSE}
grViz('
graph cycle {
 graph[layout = neato]
 node [shape = circle, style = filled, color = grey, label=""]

 0 [color = red]; 1; 2; 3;

  0 -- 1 -- 2 -- 3 -- 0
}
')
```



## The Automorphism Group Is C(2)

Since Z/4Z is abelian the inner automorphism are all trivial.

Let $\phi \in \mathrm{Aut{Z/4Z}}$, $\phi$ is uniquely determined by it's image of a generator of Z/4Z, for example 1.

And we have $\phi(2) = \phi(1+1) = \phi(1) + \phi(1)$

Let $n \in \mathrm(Z), g \in \mathrm{A}$ where A is any Abelian group. We write:
$$\begin{array}{ll}
  n \cdot g = g + g + \dots + g, (\mathrm{n \times}), n > 0\\
  0 \cdot g = 0_{\mathrm{A}}\\
  n \cdot g = -g - g - \dots -g, (\mathrm{-n \times}), n < 0
  \end{array}$$

We show that $g \mapsto n \cdot g$ is a group morphism $Z \rightarrow \mathrm{A}$.

For $\phi$ to be an automorphism, it musts map generators to generators so $\phi(1) \in \{1, 3\}$.

We show that each of those define an automorphism so the automorphism group is isomorphic to Z/2Z.


## Beat action of C(4)

|Beat  | 1 | 2 | 3 | 4 |
|------|---|---|---|---|
|e     | 1 | 2 | 3 | 4 |
|1     | 4 | 1 | 2 | 3 |
|2     | 3 | 4 | 1 | 2 |
|3     | 2 | 3 | 4 | 1 |

If we take the action over C(2) then there are 2^4 = 16 recordings

```{r}
g1 <- as.cycle(4:1)
```

For any beat B: |Stab(B)| . |B| = 4
So we have $|B| \in \{1, 2, 4\}$

    * If size B is 1
    
  Let f in B. Then all the elements of f are equal so f = 1111 or f = 0000.
    
    * If size B is 2 
    
  Let f be an element of size B. Stab(f) is of order 2 so Stab(f) = {0, 2}.
  That means $2 \bullet f = f$, which implies f(2) = f(0) and f(3) = f(1).
  So f is uniquely determined by f(0) and f(1). But we cannot have f(0) = f(1) otherwise we also have f(3) = f(1) = f(0) = f(2) and f has size 1. So we have at most 4 - 2 = 2 choices for f: 1 0 1 0 and 0 1 0 1. Clearly they are in the same beat:
    
```{r}
all(apprec(g1)(1, 0, 1, 0) == c(0, 1, 0, 1))
```

So there is only one beat of size 2

  * If size B is 4
  
  There are $16 - 2*1 - 1*2 = 12$ recordings of beats of size 4. So there are 3 beats of size 4.
  
    1. If cardinal of B is 1
    
  Then we can find f in B of the form 1000. Clearly it is a beat of size 4 and cardinal 1, and it is the only 1.
    
    2. If cardinal of B is 2
  Let f in B. We can always find a representant of the same beat starting with a 1 so we can assume f is one of:
  1 1 0 0  
  1 0 1 0 of size 2  
  1 0 0 1 = g3 . 1100
    
```{r}
all(apprec(g1^3)(1, 1, 0, 0) == c(1, 0, 0, 1))
```

So B is the beat 1 1 0 0 and we verify by trying all the possibilities that it is of size 4.

    3. If cardinal of B is 3
    
  We have a bijection with Beats of size 4 and cardinal 3 and the beats of size 4 and cardinal 1, given by $\eta$. So there is only 1 beat of size 4 cardinal 3 given by
  1110.
    
  * Conclusion
  
| Beat size | Number |
|-----------|--------|
| 1         | 2      |
| 2         | 1      |
| 4         | 3      |

| Beat | Size |
|------|------|
| 1111 | 1    |
| 0000 | 1    |
| 1010 | 2    |
| 1000 | 4    |
| 1100 | 4    |
| 1110 | 4    |

## Other beat action of C(4)

We have $C(1) \lhd C(2) \lhd C(4)$ With two C(2) composition factors

|Beat  | 1 | 2 | 3 | 4 |
|------|---|---|---|---|
|e     | 1 | 2 | 3 | 4 |
|2     | 2 | 1 | 4 | 3 |
|1     | 3 | 4 | 2 | 1 |
|3     | 4 | 3 | 1 | 2 |


This is an action because
```{r}
g <- as.word(c(3, 4, 2, 1))
g^4 == as.word(c(1:4))
g^2 == as.word(c(2, 1, 4, 3))
g^3 == as.word(c(4, 3, 1, 2))
```

This action is equivalent to the left regular action
```{r}
ord <- as.word(c(1, 3, 4, 2))
g^ord == as.word(c(4, 1, 2, 3))
```

| Beat | Size | Friends             |
|------|------|---------------------|
| 0000 | 1    | 0000                |
| 1000 | 4    | 1000 0100 0001 0010 |
| 1010 | 4    | 1010 0101 1001 0110 |
| 1100 | 2    | 1100 0011           |
| 1110 | 4    | 1110 1011 1101 0111 |
| 1111 | 1    | 1111                |

For comparison this is the beat table for C(2)xC(2)   

| Beat | Size | Friends             |
|------|------|---------------------|
| 0000 | 1    | 0000                |
| 1000 | 4    | 1000 0100 0010 0001 |
| 1010 | 2    | 1010 0101           |
| 1100 | 2    | 1100 0011           |
| 1001 | 2    | 1001 0110           |
| 1011 | 4    | 1011 0111 1110 1101 |
| 1111 | 1    | 1111                |

### What are the C(2) invariants on this action?


Here we are talking about the C(2) valued invariants. That is functions $Rec \rightarrow C(2)$ that are constant on beat actions. We call them Inv(Rec, C(2)).

We have by definition 
$$f \in Inv(Rec, C(2)) \iff \forall g \in C(4), r \in Rec, f(g \bullet r) = f(r)$$

Inv(Rec, C(2)) is an Abelian Group, subgroup of Rec


$(f_1+f_2)(g \bullet r) = f_1(g \bullet r) + f_2(g \bullet r) = f_1(r) + f_2(r) = (f_1+f_2)(r)$

Rec has $2^4 = 16$ elements. So the set Hom(Rec, C(2)) has 2^16 = 65536 elements. We can deduce that Inv(Rec, C(2)) has dimension 1, 2, 4, 8, 2^4, ... 2^16.
Inv(Rec, C(2)) cannot have 2^16 elements because the map $r \mapsto r_1$ is not a beat invariant.


For $cons: r \mapsto 1$ called constant function is a beat invariant.

The map $sum: r \mapsto \sum_{j \in 1...4} r_j$ is a beat invariant.

The map $dif: r \mapsto r_1 \times r_2 + r_3 \times r_4$ is a beat invariant.
Indeed $dif(1 \bullet r) = r_4 \times r_3 + r_1 \times r_2 = dif(r)$

\<cons, sum, dif\> est un sous-group Abelien de Inv(Rec, C(2)) d'ordre 8. 


| Beat | Size | Friends             | sum | dif |
|------|------|---------------------|-----|-----|
| 0000 | 1    | 0000                | 0   | 0   |
| 1111 | 1    | 1111                | 0   | 0   |
| 1000 | 4    | 1000 0100 0001 0010 | 1   | 0   |
| 1010 | 4    | 1010 0101 1001 0110 | 0   | 0   |
| 1100 | 2    | 1100 0011           | 0   | 1   |
| 1110 | 4    | 1110 1011 1101 0111 | 1   | 1   |

We notice that if we except the 2 constant beats, it is possible to decide to which beat belongs any recording just by computing this 2 C(2)-invariants.

One question we might ask is: if I know the Inv(Rec, C(2), {0, 2}) can I deduce the invariants for C(4)? Especially since {0, 2} is normal in C(4).

Any C(4) invariant is also a C(2) invariant

$ Inv(C(2)^4, C(4))  \subset  Inv(C(2)^4, \{0, 2\}) $

That means $Inv(C(2)^4, \{0, 2\})$ has dimension 8, 16, 2^5, ..., 2^15

This is the sub-action of {0, 2}

|Beat  | 1 | 2 | 3 | 4 |
|------|---|---|---|---|
|e     | 1 | 2 | 3 | 4 |
|2     | 2 | 1 | 4 | 3 |

We notice that this action of {0, 2} on C(2)^4 is isomorphic to the direct product of 2 actions of C(2) on C(2)^2 given by

|Beat  | 1 | 2 |
|------|---|---|
|0     | 1 | 2 |
|1     | 2 | 1 |

This space C(2)^2 has 4 elements. So Hom(C(2)^2, C(2)) has 2^4 = 16 elements
The invariants Inv(C(2)^2, C(2)) has dimension 1, 2, 4 or 8 because clearly not all the maps are invariants.

The constant map is invariant. The sum map is an invariant. The product map is also an invariant.

| Beat | Size | Full Orbit | sum | prod |
|------|------|------------|-----|------|
| 00   | 1    | 00         | 0   | 0    |
| 11   | 1    | 11         | 0   | 1    |
| 10   | 2    | 10 01      | 1   | 0    |

So we notice that the maps sum, prod and cons are independant (they generate subgroups in internal direct product). So we have dim(\<sum, prod, cons\>) = 8 Since
$$<sum, prod, cons> \subset Inv(C(2)^2, C(2))$$ and the invariants cannot have dimension 16 we have: 
$$<sum, prod, cons> = Inv(C(2)^2, C(2))$$

Now we want to find Inv(C(2)^4, {0, 2})

We notice that for f1, f2 in Inv(C(2)^2, C(2)) and f any function $C(2)^2 \rightarrow C(2)$
the map $(a, b, c, d) \mapsto f(f_1(a, b), f_2(c, d))$ is an invariant

Let F be an invariant of Inv(C(2)^4, {0, 2}). Then for every $(c, d) \in C(2)^2$ the map $(a, b) \mapsto F(a,b,c,d)$ is a Inv(C(2)^2, {0, 2}) invariant.

In other word there is mapping
$$\mathrm{copy}: Inv(C_2^4, C_2) \rightarrow (C_2^2 \rightarrow Inv(C_2^2,C_2))$$

This map is injective. Let $F_1, F_2 \in Inv(C_2^4, C_2)$ so that there exists a,b,c,d
$F_1(a,b,c,d) \neq F_2(a,b,c,d)$ which means $copy(F_1)(c, d)(a,b) \neq copy(F_2)(c, d)(a, b)$
which means $copy(F_1)(c, d) \neq copy(F_2)(c, d)$ and finally $copy(F_1) \neq copy(F_2)$


Since $C_2^2 \rightarrow Inv(C_2^2,C_2)$ has dimension $8^4 = 2^{12} = 4096$
We deduce that $|Inv(C_2^2,C_2)| \leq 2^{12}$

In particular we have $$2^3 \leq |Inv(C_2^2,C_2)| \leq 2^{12}$$

Is the copy map surjective?

Let $T: C_2^2 \rightarrow Inv(C_2^2,C_2)$

Let $G: C_2^4 \rightarrow C_2$ be defined by
$$ G(a, b, c, d) = T(c, d)(a, b)$$

G is not necessarily invariant by C(2) because nothing says that T has to be symmetric in (c, d).


## Action on 2 bits

Let A = {00, 01, 10, 11}

We define an action

|Beat  | 00 | 01 | 10 | 11 |
|------|----|----|----|----|
|0     | 00 | 01 | 10 | 11 |
|1     | 01 | 10 | 11 | 00 |
|2     | 10 | 11 | 00 | 01 |
|3     | 11 | 00 | 01 | 10 |

Projection on each sub
Let's check that it is indeed an action
```{r}
g2b_1 <- as.word(c(2, 3, 4, 1))
g2b_1^4 == as.word(1:4)
g2b_1^2 == as.word(c(3, 4, 1, 2))
g2b_1^3 == as.word(c(3, 4, 1, 2))
```

