---
title: "Z/4Z"
author: "Manu"
date: "3 March 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Group defintion

Table of the group:

| Z/4Z | 0 | 1 | 2 | 3 |
|------|---|---|---|---|
| 0    | 0 | 1 | 2 | 3 |
| 1    | 1 | 2 | 3 | 0 |
| 2    | 2 | 3 | 0 | 1 |
| 3    | 3 | 0 | 1 | 2 |

Putting the subgroup {0, 2} first:

| Z/4Z | 0 | 2 | 1 | 3 |
|------|---|---|---|---|
| 0    | 0 | 2 | 1 | 3 |
| 2    | 2 | 0 | 3 | 1 |
| 1    | 1 | 3 | 2 | 0 |
| 3    | 3 | 1 | 0 | 2 |





## Subgroups

```{r subgroup-lattice, echo = FALSE}
grViz('
graph cycle {
 graph[layout = neato]
 node [shape = circle, style = filled, color = grey]

 0 [label = "0"]; C4 [label = "C(4)"]; 2 [label = "{0, 2}"];

  0 -- 2 -- C4
}
')
```











## Graphs

```{r cycle-graph, echo = FALSE}
grViz('
graph cycle {
 graph[layout = neato]
 node [shape = circle, style = filled, color = grey, label=""]

 0 [color = red]; 1; 2; 3;

  0 -- 1 -- 2 -- 3 -- 0
}
')
```



## Automorphism group

Since Z/4Z is abelian the inner automorphism are all trivial.

Let $\phi \in \mathrm{Aut{Z/4Z}}$, $\phi$ is uniquely determined by it's image of a generator of Z/4Z, for example 1.

And we have $\phi(2) = \phi(1+1) = \phi(1) + \phi(1)$

Let $n \in \mathrm(Z), g \in \mathrm{A}$ where A is any Abelian group. We write:
$$\begin{array}{ll}
  n \cdot g = g + g + \dots + g, (\mathrm{n \times}), n > 0\\
  0 \cdot g = 0_{\mathrm{A}}\\
  n \cdot g = -g - g - \dots -g, (\mathrm{-n \times}), n < 0
  \end{array}$$

We show that $g \mapsto n \cdot g$ is a group morphism $Z \rightarrow \mathrm{A}$.

For $\phi$ to be an automorphism, it musts map generators to generators so $\phi(1) \in \{1, 3\}$.

We show that each of those define an automorphism so the automorphism group is isomorphic to Z/2Z.











## Beat action of C(4)

|Beat  | 1 | 2 | 3 | 4 |
|------|---|---|---|---|
|e     | 1 | 2 | 3 | 4 |
|1     | 4 | 1 | 2 | 3 |
|2     | 3 | 4 | 1 | 2 |
|3     | 2 | 3 | 4 | 1 |

If we take the action over C(2) then there are 2^4 = 16 recordings

```{r}
g1 <- as.cycle(4:1)
```

For any beat B: |Stab(B)| . |B| = 4
So we have $|B| \in \{1, 2, 4\}$

    * If size B is 1
    
    Let f in B. Then all the elements of f are equal so f = 1111 or f = 0000.
    
    * If size B is 2 
    
    Let f be an element of size B. Stab(f) is of order 2 so Stab(f) = {0, 2}.
    That means $2 \bullet f = f$, which implies f(2) = f(0) and f(3) = f(1).
    So f is uniquely determined by f(0) and f(1). But we cannot have f(0) = f(1) otherwise we also have f(3) = f(1) = f(0) = f(2) and f has size 1. So we have at most 4 - 2 = 2 choices for f: 1 0 1 0 and 0 1 0 1. Clearly they are in the same beat:
    
```{r}
all(apprec(g1)(1, 0, 1, 0) == c(0, 1, 0, 1))
```

So there is only one beat of size 2

  * If size B is 4
  
  There are $16 - 2*1 - 1*2 = 12$ recordings of beats of size 4. So there are 3 beats of size 4.
  
    1. If cardinal of B is 1
    Then we can find f in B of the form 1000. Clearly it is a beat of size 4 and cardinal 1, and it is the only 1.
    
    2. If cardinal of B is 2
    Let f in B. We can always find a representant of the same beat starting with a 1 so we can assume f is one of:
    1 1 0 0  
    1 0 1 0 of size 2  
    1 0 0 1 = g3 . 1100
    
```{r}
all(apprec(g1^3)(1, 1, 0, 0) == c(1, 0, 0, 1))
```

So B is the beat 1 1 0 0 and we verify by trying all the possibilities that it is of size 4.

    3. If cardinal of B is 3
    
    We have a bijection with Beats of size 4 and cardinal 3 and the beats of size 4 and cardinal 1, given by $\eta$. So there is only 1 beat of size 4 cardinal 3 given by
    1110.
    
  * Conclusion
  
| Beat size | Number |
|-----------|--------|
| 1         | 2      |
| 2         | 1      |
| 4         | 3      |

| Beat | Size |
|------|------|
| 1111 | 1    |
| 0000 | 1    |
| 1010 | 2    |
| 1000 | 4    |
| 1100 | 4    |
| 1110 | 4    |
