---
title: "1D affine transforms"
author: "Emmanuel Blonkowski"
date: "9/18/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(expm)
```

# Group of Orientation Preserving 1D Affine Transforms

The set of all functions of the form:
$$f: x \mapsto a\cdot x + b, \text{where }a \in \mathbb{R}_+^*, b \in \mathbb{R}$$

We can show that:
    - This is a group
    - It is 2 dimensional
    - Translations $x \mapsto x + b$ form a normal subgroup isomorphic to
        $\mathbb{R}$
    - Scale transformations $x \mapsto a \cdot x$ for $a \gt 0$ form a subgroup
         isomorph to $(\mathbb{R^*_+},\times)$
    - Scale transformations and translations intersects trivialy
    - Together they generate the affine transforms:
         $f:x \mapsto a\cdot x + b = D(b) \circ S(a) (x)$
    - Finally the scales transform break into a semi-direct product:
        $S \ltimes D$

# Representation As A Matrix Group

Let's have a look at the 2 matrices below
```{r}
S = matrix(c(exp(1),0,0,1), nrow = 2)
D = matrix(c(1,0,1,1), nrow = 2)
print('S')
print(S)
print('D')
D
```

## One Dimensional Subgroups of the Generators

**For D**:

1 dimensional subgroup generated by D is all the matrices of the form:
$$D^b = \begin{bmatrix}
1 & b\\
0 & 1
\end{bmatrix}$$

```{r}
D%^%10
```

We verify below that 
$$\mathcal{d} = \begin{bmatrix}
0 & 1\\
0 & 0
\end{bmatrix}$$


Is the associated element of the Lie algebra.

```{r}
d = matrix(c(0,0,1,0), 2)
print(expm(50*d))
```

**For S:**

We have:
$$S^t = \begin{bmatrix}
e^t & 0 \\ 0 & 1
\end{bmatrix}$$

We verify that $$\mathcal{s} = \begin{bmatrix}
1 & 0 \\ 0 & 0
\end{bmatrix}$$

Is the associated element of the Lie algebra.

```{r}
s = matrix(c(1, 0, 0, 0), 2)
expm(s)
```


## Commutation Relations

We check that the subgroup generated by D is normal.

$$SDS^{-1} = D^{e}$$

```{r}
S%*%(D%^%10)%*%solve(S)
```

End we can deduce that the matrix group is a semi-direct product, of the subgroup generated by the S matrix acting on the subgroup generated by the D matrix.

## Elements of the Matrix Group

From the semi-direct product decomposition we can deduce that all element of the group can be written as:

$$g = D^dS^l$$
The group product writes:
$$D^{d_1}S^{l_1}D^{d_2}S^{l_2} = D^{d_1}S^{l_1}D^{d_2}S^{-l_1}S^{l_1}S^{l_2} =\\
D^{d_1}D^{l_1d_2}S^{l_1+l_2} = D^{d_1 + l_1d_2}S^{l_1+l_2}$$


