---
title: "C(4)xC(2)"
author: "Emmanuel Blonkowski"
date: "23 May 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(DiagrammeR)
```

## Definition

We write it additively as the product $C(4) \times C(2)$.

Example of element: (3, 1)

It is a commutative group with 8 elements.

Table of elements

|Element | Order |
|--------|-------|
|(0,0)   |1      |
|(1,0)   |4      |
|(2,0)   |2      |
|(3,0)   |4      |
|(0,1)   |2      |
|(1,1)   |4      |
|(2,1)   |2      |
|(3,1)   |4      |

It has 4 elements of order 4, 3 elements of order 2.

## Cycle Graph

```{r echo = FALSE}
grViz('
graph cycle {
 graph[layout = neato]
 node [shape = circle, style = filled, color = grey, label=""]

 00 [color = red]

  00 -- 10  -- 20 -- 30  -- 00
  edge [color = "green"]
  00 -- 11  -- 20 -- 31 -- 00
  edge [color = "blue"]
  00 -- 01
  00 -- 21
}
')
```

## Subgroups

It has 2 subgroups of order 4: \<10\> and \<11\>    
It has 3 subgroups of order 2: \<01\>, \<20\> and \<21\>

Lattice of subgroups:

```{r echo = FALSE}
grViz('
graph cycle {
 graph[layout = dot]
 node [shape = circle, style = filled, color = grey]

 00 [color = red, label = "00"]; 10 [label = " 10 "]; 11 [label = " 11 "]
 G [label = "C(4)xC(2)"]

 00 -- 01
 00 -- 20
 00 -- 21

 20 -- 10 -- G
 20 -- 11 -- G
 01 -- G
 21 -- G

  {rank = same; 20; 01; 21;}
}
')
```

## Quotients

Quotient by any subgroup of order 4 gives C(2).   
Quotient by 01 yields a subgroup of order 4. We note that the class of 10 has order 4 in the quotient so $C(4) \times C(2) \ / <01> \  \sim C(4)$.   
Quotient by 20 yields a subgroup of order 4. All elements of order 4 in G are mapped to elements of order 2 in 20. So $G / 20$ has no element of order 4 and $G / 20 \sim C(2)^2$
Quotient by 21 yields a subgroup of order 4. We note that the class of 10 has order 4 in the quotient so $C(4) \times C(2) \ / <21> \  \sim C(4)$.     

**Conclusion**: there are 3 different quotients $C(2), C(4), C(2)^2$.
