---
title: "D4"
author: "Emmanuel Blonkowski"
date: "23 May 2019"
output: 
  html_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(DiagrammeR)
```

## Definition

We use the following presentation:
$$ <r,s / r^4, s^2, srs = r^3>$$

**Property** It is a non-commutative group with 8 elements.$D4 \sim C(2) \rtimes C(4)$ where C(2) acts by inverting elements of the C(4) normal subgroup.

**Proof** 
>Clearly \<r\> has order 4. Furthermore $r<r>r^{-1} = s<r>s = <r>$ and r,s generate D4 so \<r\> is normal in D4. We have :
$$0 \rightarrow <r> \rightarrow D4 \rightarrow D4/<r> \rightarrow 0$$
D4/\<r\> is generated by [s] and thus has order 2. The map that sends [s] to s right splits the above sequence. Thus $D4 ~ C(2) \rtimes C(4)$ and in particular it has 8 elements.


Table of elements

|Element | Order |
|--------|-------|
|e       |1      |
|r       |4      |
|r2      |2      |
|r3      |4      |
|s       |2      |
|sr      |2      |
|sr2     |2      |
|sr3     |2      |


It has 2 elements of order 4, 5 elements of order 2.

## Conjugacy

We have 5 conjugacy classes

- e
- r, $r^3$
- s, $sr^2$
- r2
- sr, $sr^3$

Written in tree notation:


taking s = 1(00), r = 1(10)
then r2 = 0(11), sr2 = 1(11), sr = 0(10)

- Identity 0(00)
- Center 0(11)
- 0(10), 0(01)
- s-conjugacy 1(00), 1(11)
- 4-cycle 1(10), 1(01)

Representants: 0(00), 0(11), 0(10), 1(00), 1(10)

## Cycle Graph

```{r echo = FALSE}
grViz('
graph cycle {
 graph[layout = neato]
 node [shape = circle, style = filled, color = grey, label=""]

 00 [color = red]; r [color = black]; r3 [color = black]; r2 [color = blue]
 s [color = brown]; sr2 [color = brown]
  00 -- r  -- r2 -- r3 -- 00
  edge [color = "green"]
  00 -- sr
  00 -- sr2
  00 -- sr3
  00 -- s
}
')
```

## Center of D4

The center of D4 is $<r^2> \ \sim C(2)$

## Commutator Subgroup

$srsr^3 = r^3r^3 = r^2$

## Subgroups

It has 1 subgroup ~ C(4): \<r\>    
It has 5 subgroups of order 2: $<r^2>$, \<s\>, \<sr\>, $<sr^2>$ and $<sr^3>$    
It has 2 subgroups ~ C(2)xC(2): $<r^2, sr^2>$, $<r^2, sr>$,

**Property** The only subgroups of D4 isomorph to C(2)xC(2) are: $<r^2, s>$ and $<r^2, sr>$.

**Proof** 
> Any subgroup isomorph to C(2)xC(2) can be generated by 2 elements of order 2 that commutes together. D4 has 5 elements of order 2, below we list how they commute with each other:

>
$<r^2>$ commutes with all elements
\<s\> commutes with $<r^2>$, $<sr^2>$
\<sr\> commutes with $<r^2>$, $<sr^3>$
$<sr^2>$ commutes with $<r^2>$, \<s\>
$<sr^3>$ commutes with $<r^2>$, \<sr\>

>
For each pair of commutating elements of order 2 we have a subgroup isomorph to C(2)xC(2). Accordingly the possible candidates are:
$<r^2, s>$, $<r^2, sr>$, $<r^2, sr^2>$, $<r^2, sr^3>$, 
$<s, sr^2>$, $<sr, sr^3>$

>
Some of these groups are actually the same. Removing the groups that are the same:
$<r^2, s>$, $<r^2, sr>$


Lattice of subgroups:

```{r echo = FALSE}
grViz('
graph cycle {
 graph[layout = dot]
 node [shape = circle, style = filled, color = grey]

 e [color = red, label = "e"]; r [label = "  r   ", color = "green"]; 
 r2 [color = "blue", label = " r2 "]; s [label = " s "];
  sr [label = " sr"];
 r2_s [label = "r2, s", color = "green"];  r2_sr [label = "r2, sr", color = "green"];
 G [label = "  D4  "]

 e -- r2 -- r -- G
 e -- s -- r2_s -- G
 e -- sr -- r2_sr -- G
 e -- sr2 -- r2_s
 e -- sr3 -- r2_sr

 r2 -- r2_s
 r2 -- r2_sr

}
')
```

## Normal Subgroups

We check that r, r2, \<r2, s\>, <\r2, sr\> are the only normal subgroups of D4.

```{r echo = FALSE}
grViz('
graph cycle {
 graph[layout = dot]
 node [shape = circle, style = filled, color = grey]

 e [color = red, label = "e"]; r [label = "  r   ", color = "green"]; 
 r2 [color = "blue", label = " r2 "]; 
 r2_s [label = "r2, s", color = "green"];  r2_sr [label = "r2, sr", color = "green"];
 G [label = "  D4  "]

 e -- r2 -- r -- G
 r2 -- r2_s -- G
 r2 -- r2_sr -- G

}
')
```

## Quotients

3 different quotients onto C(2):
$$0 \rightarrow r \rightarrow D4 \rightarrow C(2) \rightarrow 0$$
$$0 \rightarrow <r^2, s> \rightarrow D4 \rightarrow C(2) \rightarrow 0$$
$$0 \rightarrow <r^2, sr> \rightarrow D4 \rightarrow C(2) \rightarrow 0$$

The quotient $D4 /r^2$ cannot have any element of order 4, further it is itself a group of order 4, we conclude that $D4 / r^2 \sim C(2) \times C(2)$

Thus we have one last quotient:

$$0 \rightarrow <r^2> \rightarrow D4 \rightarrow C(2)^2 \rightarrow 0$$
I don't think this is right-split.

## Automorphisms

### Inner Automorphisms

The group of inner automorphisms is given by $D_4 / Z = D_4/r^2 \sim C(2) \times C(2)$

It contains:

- Identity. $(r, s)$
- s-type = $(r\ r^3)(sr\ sr^3)$, s, r2 is fixed. $(r^3, s)$
- r-type = $(s\ sr^2)(sr\ sr^3)$ (whole \<r\> is fixed). $(r, sr^2)$
- sr-type = $(r\ r^3)(s\ sr^2)$, sr, r2 fixed. $(r^3, sr^2)$

### Full Automorphism Group

Since D4 is generated by r, s an auto-morphism $\phi$ is uniquely defined by the image $\phi(r)$ and $\phi(s)$.


Considering the orders:
- $\phi(r) \in \{r, r^3\}$
- $\phi(s) \in \{s, sr, sr^2, sr^3\}$ Note that we can remove $r^2$ from this list because it belongs to the center, and $\phi$ maps the center to the center.


**1st** let's assume $\phi(r) = r$
We have $\phi(srs) = \phi(s)r\phi(s)$ but also $\phi(srs) = \phi(r^3) = r^3$
If $\phi(s) \in \{s, sr^2\}$ then we find an inner-automorphism.
So we can assume either $\phi(s) = sr$ or $\phi(s) = sr^3$

- If $\phi(s) = sr$ then $\phi(srs) = srrsr = sr^2sr = r^3$. So this condition is 
verified. This automorphism is given by $(s\ sr\ sr2\ sr3)$ (fixes \<r\>)
- If $\phi(s) = sr^3$ then $\phi(srs) = sr^3rsr^3 = ssr^3 = r^3$. So this condition is also verified. This automorphism is given by $(s\ sr^3\ sr^2\ sr)$ (fixes \<r\>)

**2nd** let's assume $\phi(r) = r^3$
We have $\phi(srs) = \phi(s)r^3\phi(s)$ but also $\phi(srs) = \phi(r^3) = r$
If $\phi(s) \in \{s, sr^2\}$ then we find an inner-automorphism.

- If $\phi(s) = sr$ then $\phi(srs) = srr^3sr = ssr = r$. So this works.
We get $(r\ r^3)(s\ sr\ sr^2\ sr^3)$
- If $\phi(s) = sr^3$ we get $(r\ r^3)(s\ sr\ sr^2\ sr^3)$

So we have a bijection:
$$\{r, r^3\} \times \{s, sr, sr^2, sr^3\} \rightarrow \text{Aut}(D_4)$$


### Structure of Automorphism Group

**Property** A is generated by $(r, sr)$, $(r^3, s)$ and $(r, sr^2)$.

> Proof:
>
$Inn(D_4)$ is normal in this group. Which means the full automorphism group A admits a short exact sequence: 
$$0 \rightarrow \text{Inn} \rightarrow A \rightarrow \text{Out} \rightarrow 0$$
>
So Out is generated by $(r, sr)$. Inn is generated by $(r^3, s)$ and $(r, sr^2)$.


**Property** A is isomorph to D4.

> Proof
We'll check that we have the same presentation:
>
- $(r, sr)$ has order 4
- $(r^3, s)$ has order 2
- $(r^3, s) \circ (r, sr) \circ (r^3, s) = (r^3, sr^3)$
  Since $r \mapsto r^3 \mapsto r^3r^3r^3 = r \mapsto r^3$ and
  $s \mapsto s \mapsto sr \mapsto sr^3$
  


