---
title: "S3xC(2)"
author: "Manu"
date: "3 August 2018"
output: 
  html_document:
    toc: true
    number_sections: true
    css: ../custom.css
---

```{r setup, include=FALSE}
library(permutations)
library(DiagrammeR)
source('../utils.R')
knitr::opts_chunk$set(echo = TRUE)
```

# S3xC(2)

S3xC(2) is the direct product of the symmetric group on 3 elements by C(2), the cyclic group of order 2. S3xC(2) is a group of order 12. This group is interesting because it can both be written has a direct product and as a non-trivial semi-direct product.

Since S3 admits $<r^3, s^2, srsr>$ as a presentation
S3xC(2) admits $<r^3, s^2, i^2, srsr, is = si, ir = ri>$, where i is $(e_{S3}, 1)$

## Order of the Elements

| Elements | e | r | r2 | s | sr | sr2 | i | ir | ir2 | is | isr | isr2 |
|----------|---|---|----|---|----|-----|---|----|-----|----|-----|------|
| Order    | 1 | 3 | 3  | 2 | 2  | 2   | 2 | 6  | 6   | 2  | 2   | 2    |

The **conjugacy classes** are:

1.  {e}
2.  {r, r2}
3.  {ir, ir2}
4.  {i}
5.  {s}
6.  {sr, sr2}
7.  {is}
8.  {isr, isr2}

Below is the **cycle graph**. We used the same color for the nodes which are conjugates. The nodes in grey are alone in their conjugacy class.
```{r echo = FALSE}
grViz('
graph cycle {
 graph[layout = neato]
 node [shape = circle, style = filled, color = grey, label=""]

  e [color = red]; r [color = green]; r2 [color = green]; 
 ir [color = blue]; ir2 [color = blue];
  isr [color = brown]; isr2 [color = brown]
  sr [color = yellow]; sr2 [color = yellow]; i; 
  is; 


  e -- ir -- r2 -- i -- r -- ir2 -- e
  e -- isr
  e -- isr2
  e -- sr
  e -- sr2
  e -- is
  e -- s

}')
```

## Normal Decomposition

$$0 \lhd \ <r> \ \lhd S3 \lhd S3 \times C(2)$$

With decomposition factors isomorphic to: \<r\>, \<s\>, \<i\>

Another isomorphic way to write this chain:

$$0 \lhd C(3) \lhd C(3) \rtimes C(2) \lhd [\lhd C(3) \rtimes C(2)] \times C(2)$$

With decomposition factors isomorphic to: C(3), C(2), C(2)

## Normal Subgroups

\<r\> is the only **normal subgroups contained in S3** that is normal in the full group.
It is a minimal normal subgroup in that there is no non-trivial subgroup of \<r\> that is also normal in the full group.

**Property** S3xC(2)/\<r\> is isomorphic to $C(2) \times C(2)$

> G/r is generated by \<s, i / si = is\>. We can check easily that there is no g in \<r\> such that sg = i, so that s and i are not projected to the same value in G/r.
So G/r is isomorphic to C(2)xC(2).

**Property** The extension $0 \rightarrow <r> \rightarrow S3\times C(2) \rightarrow C(2) \times C(2) \rightarrow 0$, is right split but not left-split.

> Proof: The application that sends [s] to s and [i] to i right splits the sequence. THe action morphism optained has a non-inner action on \<r\>, so the sequence cannot be left-split.

The **complete list of normal subgroups** is {e}, \<r\>, \<i\>, \<r, i\>, \<r, is\>, \<r, s\> and S3xC(2).

> Let H be a normal subgroup of the full group. Normal subroups are closed under intersection so $H \cap <r>$ is also a normal subgroup. So we know $H \cap <r> = {e}$ pr $H \cap <r> = <r>$.
>
1. If $H \cap <r> = <r>$ then H = \<r\>, S3x{0}, \<r, i\>, \<r, is\>, \<r, s\> or S3xC(2)
>
2. If $H \cap <r> = {e}$. Then H\<r\> is a normal subgroup containing \<r\> so it is one of the groups above: \<r\>, S3x{0}, \<r, i\>, \<r, is\> or S3xC(2).
We also get a split exact sequence 
$$0 \rightarrow <r> \rightarrow H<r> \rightarrow H \rightarrow 0$$
If H\<r\> is \<r\> then H = {e}.
If H\<r\> is S3 it cannot be because there are no split sequences starting with $0 \rightarrow <r> \rightarrow S3$ so it is not possible.
Same argument if H\<r\> is \<r, is\>, it cannot be.
If H\<r\> is \<r, i\> then H = \<i\>.
If H\<r\> is S3xC(2) it also cannot be because there are no split sequences starting with $0 \rightarrow <r> \rightarrow S3xC(2)$.

Here is the **normal lattice**, rs and ris are both isomorphic to S3, all the other groups are distinct.

```{r normal-lattice, echo = FALSE}
grViz('
digraph boxes_and_circles {

  graph [overlap = true, fontsize = 10, rankdir = "BT"]

  node [shape = circle,
        fontname = Helvetica,
        fillcolor = cyan,
        style = filled]
  e; r [width = 0.75, fontsize = 16]; i [width = 0.62, fontsize = 16];
  node [width = 1, fontsize = 18]
  rs [fillcolor = darkorchid2] ; ri ; ris [fillcolor = darkorchid2] ;
  S3xC2 [width = 2, fontsize = 22]

  e->r->rs->S3xC2
  r->ris->S3xC2
  e->i->ri->S3xC2
  r->ri
}
')
```

# S3xC(2) as a Semi-Direct Product

Let's write $H = \{e, is\}$. H is a subgroup of S3xC(2) of order  2.
We have

1. $S3 \lhd S3 \times C(2)$
2. $S3 \cap H = \{e\}$
3. $S3 \times C(2) = S3H = HS3$

So S3xC(2) is a semi-direct product $S3 \rtimes H$, with the action of H onto S3 given by
$$isris = r^2 \ \& \ issis = s$$
In particular the action is not trivial.

